Future challenges for Research Software Engineering

### 2025-05-19 future_challenges_for_RSEng.tex
add some more of the original text
rewrite story. Kyle gone.

The lessons learned from Human Factors research of the previous decades \cite{Simkute2025} applied to AI tools helped resolve the initial issues outlined above.
We envision that AI will be transforming research processes by enabling rapid analysis of massive datasets, uncovering patterns and insights that would have been impossible or prohibitively time-consuming for humans alone to find.
In research and development, AI will accelerate scientific breakthroughs in fields like healthcare, climate science, and materials engineering by automating complex modelling and predictive tasks.

@@

\begin{story}{The greater world of Kim and Kay}{ethics}
10 years have passed since Kim and Kay started on their journey of being RSEs. Back then,
at the universities of Eden and Utopia, life was more orderly and harmonious. Kay is now working on a project that develops
software for cheap water monitoring devices in the Global South. Kim is currently designing the software service stack
at the university, and increasingly struggles to find compelling Open Source alternatives that enables them keep complete control of their data
rather than relying on software and tools that are backed by large corporate entities. So far she still has the support of the university board, since it values
strategic autonomy.
Her son Kyle on the other hand started out with a degree in Physics and then took one of the newly created RSE Master's courses,
but quickly decided that work in academia with its time limited contracts is not attractive.
Governments around the world are struggling to reposition themselves in this emerging de-globalised world with its new blocks
and therefore have de-prioritised publicly funded research. Kyle decided to work for a defence company since they offer
more secure employment in this polarised world.
\end{story}

>>>>

The lessons learned from Human Factors research of the previous decades \cite{Simkute2025} applied to AI tools helped resolve the initial issues outlined above.
We envision that AI will be transforming research processes by enabling rapid analysis of massive datasets, uncovering patterns and insights that would have been impossible or prohibitively time-consuming for humans alone to find.
In research and development, AI will accelerate scientific breakthroughs in fields like healthcare, climate science, and materials engineering by automating complex modelling and predictive tasks.
Developers will be shifting from manual coding to orchestrating AI-driven workflows, focusing more on high-level design, integration, and creative problem-solving while AI handles repetitive or routine tasks.

@@

\begin{story}{The greater world of Kim and Kay}{ethics}
10 years have passed since Kim and Kay started on their journey of being RSEs. Back then,
at the universities of Eden and Utopia, life was more orderly and harmonious.
Kim is currently designing the software service stack at the university.
She increasingly struggles to find compelling Open Source alternatives that enables them to keep complete control of their data
rather than relying on software and tools that are backed by large, foreign corporate entities.
So far, she still has the support of the university board, since it values strategic autonomy.
Kay used to work on a project that developed
software for cheap water monitoring devices in the Global South, but now funding has dried up.
Having gone through another fixed term contract he is now fed up
with work in academia and decides that its time limited contracts are not attractive.
Now Governments around the world are struggling to reposition themselves in this emerging de-globalised world with its new blocks and therefore have de-prioritised publicly funded research.
Kay sees the writings on the wall and decides to work for a defence company since they offer more secure employment in this polarised world.
\end{story}
